---
title: "PracticeMidterm"
author: "Jim Liu"
date: "10/13/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
data = read.csv("/Users/jim/Dropbox (GaTech)/Courses/ISyE6402/Midterm/HouseData.csv", header=TRUE)
data.train = data[1:(nrow(data)-4),]
data.test = data[(nrow(data)-3):nrow(data),]
data.train
```


Question 1.
Provide the time series plot of the 'Houses for Sale' data. Which assumptions of stationarity, if any, seem to be violated.
Provide the time series plot of the 1st order difference of the Houses for Sale data.

```{r Q1 time series}
HFS = data.train$Houses_for_Sale
HFS.ts = ts(data=HFS, start=1976, frequency = 4)
ts.plot(HFS.ts, col='blue')
```

Constant mean/constant variance

```{r HFS Diff}
HFS.dif = diff(HFS.ts)
ts.plot(HFS.dif,main='House of Sale - Differenced',ylab='Sale Difference',col='blue')
```



Question 2.

Identify the best ARIMA model to fit the Houses for Sale data according to the AIC criterion. Use regular 1st order differencing and consider up to order 5 for AR and MA terms. Report the AIC and estimated coefficients of your selected model. Which coefficients are significant at the 5% significance level?

Provide plots of the ACF and PACF of the residuals from your ARIMA model and perform the Ljung-Box test. Do the residuals appear uncorrelated?

```{r ARIMA}
n = length(HFS.ts)
test_modelA <- function(p,d,q){
mod = arima(HFS.ts, order=c(p,d,q), method="ML")
current.aic = AIC(mod)
df = data.frame(p,d,q,current.aic)
names(df) <- c("p","d","q","AIC")
print(paste(p,d,q,current.aic,sep=" "))
return(df)
}
orders = data.frame(Inf,Inf,Inf,Inf)
names(orders) <- c("p","d","q","AIC")

for (p in 0:5){
  for (d in 0:1){
    for (q in 0:5) {
      possibleError <- tryCatch(
        orders<-rbind(orders,test_modelA(p,d,q)),
        error=function(e) e
      )
      if(inherits(possibleError, "error")) next

    }
  }
}
```

```{r ARIMA select}
orders <- orders[order(-orders$AIC),]
tail(orders)
```

```{r HOS ARIMA(5, 1, 2)}
final_HFS = arima(HFS.ts, order = c(5,1,2), method = "ML")
resids.HFS = resid(final_HFS)
## Residual Analysis
par(mfrow=c(2,2))
plot(resids.HFS, ylab='Standardized Residuals', main="Residual Plot")
abline(h=0)
acf(resids.HFS,main= 'ACF of the Model Residuals', lag.max = 52*4)
pacf(resids.HFS,main='PACF of the Model Residuals', lag.max = 52*4)
qqnorm(resids.HFS)
qqline(resids.HFS)
```

```{r hso Tests}
## Test for Independence for final model
Box.test(resids.HFS, lag = (5+1+2), type = "Box-Pierce", fitdf = (5+2))
Box.test(resids.HFS, lag = (5+1+2), type = "Ljung-Box", fitdf = (5+2))
```

```{r HSO}
library(lmtest)
coeftest(final_HFS)
```

Question 3.

Plot the forecasts for the 2017 data using the ARIMA model fit in Question 2 alongside the actual observations.
Using the precision measure, assess the prediction accuracy of the model. 

```{r EU forecasts, echo=FALSE}
n = length(data$Houses_for_Sale)
nfit = n-4
HFS.all=ts(data$Houses_for_Sale, frequency = 4, start = 1976)
outprice.EU = arima(data$Houses_for_Sale[1:nfit], order = c(5,1,2),method = "ML")
outpred.EU = predict(outprice.EU,n.ahead=4)
ubound.EU = outpred.EU$pred+1.96*outpred.EU$se
lbound.EU = outpred.EU$pred-1.96*outpred.EU$se
ymin = min(lbound.EU)
ymax = max(ubound.EU)
plot(time(HFS.all)[160:168],data$Houses_for_Sale[160:168],type="l", ylim=c(ymin,ymax), xlab="Time", ylab="Rate", main="EU Forecasting")
points(time(HFS.all)[165:168],outpred.EU$pred,col="red")
lines(time(HFS.all)[165:168],ubound.EU,lty=3,lwd= 2, col="blue")
lines(time(HFS.all)[165:168],lbound.EU,lty=3,lwd= 2, col="blue")
```

